---
title: "Jaguar Landcover of coexistence"
author: "Vanesa Bejarano Alegre"
date: "5/24/2020"
output: word_document
---
# Prepare data for jaguars - Chapter 1: Landcover of coexistence
* The purpose of this script is:
+ 1. Have the data clean and fixed

# First part : Clean

## Install and load packages

```{r Packages, message=FALSE, warning=FALSE}

rm(list= ls())  #clean all before to start

listpacks <- c("tidyverse","lubridate", "lutz", "sf", "amt")

if(!require(install.load)) install.packages('install.load'); library(install.load)
install.load::install_load(listpacks)

```

### R Packages
**_Tidyverse_** is a collection of R-packages designed for data science.

**_Lubridate_** is an R-package that makes it easier to work with dates and times. 

**_Lutz_** is an R-package that help us to find the time zones of the individuals through location.

**_sf_** is an R-package that help us to convert to UTM our data


## Set working directory and load data

```{r Directory, message=FALSE}

setwd("D:/PhD_Jaguar/Jaguar01/00_Script_to_submit/Jaguar_Data")

jaguar <- read.csv('jaguar_movement_data.csv', header = T, sep = ',', dec = '.', comment.char = '')

info <- read.csv('Jaguar_additional_information.csv', header = T, sep = ',', dec = '.', comment.char = '')

source("00_Jaguar01_Source.R")

jaguar <-  merge(jaguar, info, by.x = "individual.local.identifier..ID.", by.y = "ï..ID")

```

Next, we change the names of  long names of variables
#### First we check the current names

```{r check variables}
variable.names(jaguar)
```

```{r delete and change variables names}

#Delete some columns
jaguar <- dplyr::select(jaguar, -c("individual.taxon.canonical.name","country", "study.name","Collar.Type", "Collar.Brand","Planned.Schedule.Expected", "Observations","X..Locations","Project.Leader","Contact","Checked.plan.schedule" ))

#Change variables
jaguar <- jaguar %>% 
  rename(
    individual.local.identifier=individual.local.identifier..ID.,
    event= Event_ID,
    timestamp= timestamp,
    location.long=location.long,
    location.lat=location.lat,
    tag=tag.local.identifier,
    sex=Sex,
    age=Age,
    weight= Weight,
    country=Country,
    schedule=Planned.Schedule.Expected..h.,
    ecoregion=ECO.REGION
  )
```

*** 
#     Data preparation process
Firstly we will standardize date and time and create a new column 'timestampNEW' with this new information. Then we will change the date-time  for each time zone that the jaguar data is found.

#### Standardize date and time

```{r Date-Time, results=hide}
jaguar$timestampUTC <- mdy_hm(jaguar$timestamp)

as_datetime(jaguar$timestampUTC) #checking if everythink date-time format [ok]

any(is.na(jaguar$timestampUTC)) #cheking if we missed a data-time format [ok]

str(jaguar$timestampUTC) #[ok]

```

Then we order all data with _id_ and _timestampUTC_ and we change the date-time to the true time zone

#### Order the columns

```{r Order columns, results='hide'}
jaguar <- arrange(jaguar, individual.local.identifier, timestampUTC)
```

#### Time zones

```{r found tz,}
jaguar$tz <- tz_lookup_coords(lat= jaguar$location.lat, lon = jaguar$location.long, method = "accurate")
```

```{r timezone, }

for (i in 1:nrow(jaguar)){
  jaguar$timestampTZ[i] <- strftime(jaguar$timestampUTC[i],
         format="%Y-%m-%d %H:%M", 
tz = as.character(jaguar$tz[i]),
usetz = FALSE)
  }
  

# Order everything just in case
jaguar <- arrange(jaguar, individual.local.identifier, timestampTZ)

#transform to a POSIXct

jaguar$timestampTZ <- ymd_hm(jaguar$timestampTZ)

```


Now we will check if the dataset has duplicate data, then we will remove the collected points within a period of less than 25 minutes and the we will order it again

#### Checking duplicate data

```{r, duplicate}
Dupl<-duplicated(jaguar[,c("timestampTZ", "location.lat","location.long","individual.local.identifier")])
sum(Dupl) #no dumplicates
```

```{r Calculation of time between fixes,}
inds <- unique(jaguar$individual.local.identifier)
jaguar$time.diff <- unlist(sapply(inds, FUN = time.between, dat = jaguar))

```

#### Remove points less than 25 min or 1500 sec
```{r Remove point, message=FALSE, results= 'hide'}
datetimelag <- jaguar %>% 
  group_by(individual.local.identifier) %>% 
  mutate(jag_lag = timestampUTC - lag(timestampUTC, default = NA))

jag_exc <- dplyr::filter(datetimelag, jag_lag < 1500) # 25min =1500 sec
jag_remove <- anti_join(datetimelag, jag_exc)

jaguar <- jag_remove


```

#### Order the points again
```{r Order point, results='hide'}
#Data frame
jaguar <- arrange(jaguar, individual.local.identifier, timestampTZ)
jaguar$event <- seq.int(nrow(jaguar))

jaguar <- as.tibble(jaguar)
```
From here we will work only with tibble. 

***
# UTM
#### Convert to UTM
```{r UTM, results= 'hide', message= FALSE}

sf_sample <- sf::st_as_sf(jaguar, coords = c("location.long", "location.lat"),
                          crs = 4326)

jaguarUTM <- sf_sample%>%
  do(cbind(., st_coordinates(.))) %>%
  mutate(EPSG = utm_epsg(lat = Y, lon = X )) %>%
  group_by(EPSG) %>%
  do(cbind(as.data.frame(.) %>% 
             select(event, individual.local.identifier, sex, age, weight, country, ecoregion, schedule, tz, timestampUTC, timestampTZ,time.diff),
           st_coordinates(st_transform(., crs = head(.$EPSG, 1))))) %>%
  ungroup()

jaguarUTM <- arrange(jaguarUTM, individual.local.identifier, timestampTZ)

# Adding lat and lon data

jaguarUTM$location.long <- jaguar$location.long

jaguarUTM$location.lat <- jaguar$location.lat
```

*** 

## Adding new columns
Finally, to have our data sets ready, we are going to add some columns that we consider important.
First, we will add a column that tells us if it is day or night 

#### Day or night column
```{r daynight period}
jaguarUTM$hour <- hour(jaguarUTM$timestampTZ) #create a column
jaguarUTM$DayNight <-  ifelse( jaguarUTM$hour <6 | 17 < jaguarUTM$hour, "Night", "Day")
jaguarUTM$hour <- NULL #then deleted it
```


This step is to save all varibales selected.
Now we write as a __.txt__ format to use for others analysis

#### Save the work
```{r write table}

write.table(jaguarUTM ,file="00_jaguar01.txt",row.names = F,quote=F,col.names=T,sep="\t")


```

